CF Registry Analysis

### 2022-12-21 Christine Chan/CF Registry/cf_registry_analysis.qmd
python section not necessary

```{r}
#| include: false
library(tidyverse)
library(table1)
# All data
df = read.csv("Z:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Christine Chan/CF registry data/CFF19_Annualized_Del1.csv",
              na.strings = c("","N/A"))
df = df %>% arrange(eDWID,ReviewYear) 
# For those with CFRD by OGTT, find first visit with diagnosis
first_cfrd = df %>% filter(A_twohour_bloodglucose >= 200) %>% 
  group_by(eDWID) %>% slice_min(ReviewYear) %>% 
  select(eDWID,ReviewYear) %>% rename(first_cfrd = ReviewYear)
df = left_join(df,first_cfrd,by = "eDWID")
df = df %>% group_by(eDWID) %>% 
  mutate(cfrd_ogtt = case_when(
    ReviewYear < first_cfrd ~ "No",
    ReviewYear >= first_cfrd ~ "Yes"
  )) %>%
  ungroup()
df$cfrd_ogtt[is.na(df$cfrd_ogtt)] = "No"
diagnosis = df %>% 
  select(eDWID,ReviewYear,Age_YrEnd,first_cfrd,cfrd_ogtt) %>%
  group_by(eDWID) %>%
  mutate(`Age in 2018` = 2018-(ReviewYear-Age_YrEnd))
```

>>>>

```{r}
#| include: false
library(knitr)
library(tidyverse)
library(table1)
# All data
df = read.csv("Z:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Christine Chan/CF registry data/CFF19_Annualized_Del1.csv",
              na.strings = c("","N/A"))
df = df %>% arrange(eDWID,ReviewYear) %>% 
  mutate(`Age in 2018` = 2018-(ReviewYear-Age_YrEnd))
# For those with CFRD by OGTT, find first visit with diagnosis
first_cfrd = df %>% filter(A_twohour_bloodglucose >= 200) %>% 
  group_by(eDWID) %>% slice_min(ReviewYear) %>% 
  select(eDWID,ReviewYear,`Age in 2018`) %>% rename(first_cfrd = ReviewYear)
df = left_join(df,first_cfrd,by = c("eDWID","Age in 2018"))
df = df %>% group_by(eDWID) %>% 
  mutate(cfrd_ogtt = case_when(
    ReviewYear < first_cfrd ~ "No",
    ReviewYear >= first_cfrd ~ "Yes"
  )) %>%
  ungroup()
df$cfrd_ogtt[is.na(df$cfrd_ogtt)] = "No"
df = df %>% 
  select(eDWID,ReviewYear,`Age in 2018`,first_cfrd,cfrd_ogtt)
ages = df %>% group_by(eDWID) %>% 
  summarise(`Age in 2018` = names(sort(table(`Age in 2018`), decreasing = TRUE)[1]))
ages$`Age in 2018` = round(as.numeric(ages$`Age in 2018`),1)
```

```{r}
# Person-years variable
py=1000
# For each year, get the number of people diagnosed with CFRD that year
years = 2003:2018
cases = lapply(years, function(y){
  total = unique(first_cfrd$eDWID[first_cfrd$first_cfrd==y])
  cases_age_10 = total[total %in% ages$eDWID[ages$`Age in 2018` < 10]]
  cases_age_10_18 = total[total %in% ages$eDWID[ages$`Age in 2018` >= 10 & ages$`Age in 2018` < 19]]
  cases_age_19_30 = total[total %in% ages$eDWID[ages$`Age in 2018` >= 19 & ages$`Age in 2018` < 31]]
  cases_age_30 = total[total %in% ages$eDWID[ages$`Age in 2018` >= 31]]
  return(list("total"=total,"cases_age_10"=cases_age_10,
              "cases_age_10_18"=cases_age_10_18,
              "cases_age_19_30"=cases_age_19_30,
              "cases_age_30"=cases_age_30))
})
# Double check that there is no overlap between sets (there isn't!)
# lapply(1:length(cases), function(y){intersect(cases[[y]],unlist(cases[-y]))})
# For each year, get the number of people at risk
at_risk = lapply(years, function(y){
  ar = unique(df$eDWID[df$ReviewYear==y])
  ar = ar[!ar %in% unlist(cases[which(years<=y)]$total)]
  ar_age_10 = intersect(df$eDWID[df$ReviewYear==y],ages$eDWID[ages$`Age in 2018`<10])
  ar_age_10_18 = intersect(df$eDWID[df$ReviewYear==y],
                           ages$eDWID[ages$`Age in 2018`>=10 & ages$`Age in 2018`<19])
  ar_age_19_30 = intersect(df$eDWID[df$ReviewYear==y],
                           ages$eDWID[ages$`Age in 2018`>=19 & ages$`Age in 2018`<31])
  ar_age_30 = intersect(df$eDWID[df$ReviewYear==y],ages$eDWID[ages$`Age in 2018`>=31])
  return(list("total"=ar,"ar_age_10"=ar_age_10,"ar_age_10_18"=ar_age_10_18,
              "ar_age_19_30"=ar_age_19_30,"ar_age_30"=ar_age_30))
})
# Put everything in one table
inc_df = lapply(years, function(y){
  # Total
  new_cases = length(cases[[which(years==y)]]$total)
  total_at_risk = length(at_risk[[which(years==y)]]$total)
  total_inc = new_cases/(total_at_risk/py)
  # < 10
  new_cases_10 = length(cases[[which(years==y)]]$cases_age_10)
  at_risk_10 = length(at_risk[[which(years==y)]]$ar_age_10)
  inc_10 = new_cases_10/(at_risk_10/py)
  # 10 - 18
  new_cases_10_18 = length(cases[[which(years==y)]]$cases_age_10_18)
  at_risk_10_18 = length(at_risk[[which(years==y)]]$ar_age_10_18)
  inc_10_18 = new_cases_10_18/(at_risk_10_18/py)
  # 19 - 30
  new_cases_19_30 = length(cases[[which(years==y)]]$cases_age_19_30)
  at_risk_19_30 = length(at_risk[[which(years==y)]]$ar_age_19_30)
  inc_19_30 = new_cases_19_30/(at_risk_19_30/py)
  # > 30
  new_cases_30 = length(cases[[which(years==y)]]$cases_age_30)
  at_risk_30 = length(at_risk[[which(years==y)]]$ar_age_30)
  inc_30 = new_cases_30/(at_risk_30/py)
  return(list("Year" = y,
              "Total New Cases" = new_cases,
              "Total At Risk" = total_at_risk,
              "Total Incidence" = total_inc,
              "Total New Cases Age < 10" = new_cases_10,
              "Total At Risk Age < 10" = at_risk_10,
              "Total Incidence Age < 10" = inc_10,
              "Total New Cases Age 10-18" = new_cases_10_18,
              "Total At Risk Age 10-18" = at_risk_10_18,
              "Total Incidence Age 10-18" = inc_10_18,
              "Total New Cases Age 19-30" = new_cases_19_30,
              "Total At Risk Age 19-30" = at_risk_19_30,
              "Total Incidence Age 19-30" = inc_19_30,
              "Total New Cases Age > 30" = new_cases_30,
              "Total At Risk Age > 30" = at_risk_30,
              "Total Incidence Age > 30" = inc_30))
})
inc_df = do.call(rbind,inc_df)
kable(inc_df,digits=3)
```
